Well-quasi-orderings on word languages

### 2025-01-13 src/introduction.tex
Rework the intro and prelims

\AP When considering an algorithm based on a \kl{well-quasi-ordering}, the
runtime of the algorithm is deeply related to the ``complexity'' of the
underlying quasi-order~\cite{SCHMITZ17}. One way to measure this complexity is
to consider its so-called \kl{ordinal invariants}: for instance, the
\kl{maximal order type} (or m.o.t.), originally defined by De Jongh and Parikh
\cite{dejongh77}, is the order type of the maximal linearization of a
well-quasi-ordered set. In the case of a finite set, the m.o.t. is precisely
the size of the set. Better runtime bounds were obtained by considering two
other parameters~\cite{SCHMITZ19}: the \kl{ordinal height} introduced by
Schmidt \cite{schmidt81}, and the \kl{ordinal width} of Kříž and Thomas
\cite{kriz90b}. Therefore, when characterizing \kl{well-quasi-ordered}
languages, we will also be interested in deriving upper bounds on their
\kl{ordinal invariants}. We refer to \cref{ordinal-invariants:subsec} for a
more detailed introduction to these parameters and ordinal computations in
general.

\subparagraph{Contribution} In this paper, we focus on words over a finite
alphabet $\Sigma$, and characterize subsets $L \subseteq \Sigma^*$ that are
\kl{well-quasi-ordered} by the \kl{prefix relation}, the \kl{suffix relation},
and the \kl{infix relation}. Furthermore, we devise decision algorithms
whenever the languages are given by reasonable computational models. Finally,
we provide relatively low bounds on the possible \kl{ordinal invariants} of
such languages.

Specifically, for the \kl{prefix relation} and the
\kl{suffix relation}, we are able to show the following decomposition into
\intro{chains}, that are totally ordered ascending sequences.

{
\renewcommand{\labelenumi}{R\arabic{enumi}}
\begin{enumerate}
	\item A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{prefix relation} if and only if it is a finite union of \kl{chains} of the \kl{prefix relation}.
    \hfill (\cref{prefixes:thm} p.\pageref{prefixes:thm})
	\item A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{suffix relation} if and only if it is a finite union of \kl{chains} of the \kl{suffix relation}.
    \hfill (\cref{prefixes:thm} p.\pageref{prefixes:thm})
\end{enumerate}
}

Note that \kl{chains} are the simplest possible \kl{well-quasi-ordered} sets
(they are \kl{totally ordered} and \kl{well-founded}) and it is known that finite unions
of \kl{well-quasi-ordered} sets are \kl{well-quasi-ordered}. As a consequence,
the above characterization states that only the simplest possible
\kl{well-quasi-ordered} sets are \kl{well-quasi-ordered} by the \kl{prefix
relation}. Let us furthermore highlight that this characterization holds
without any restriction on the decidability of the language $L$ itself.\footnote{But
heavily relies on the assumption that $\Sigma$ is finite} This allows us to
derive the following bounds on the \kl{ordinal invariants} of such
\kl{well-quasi-ordered} languages, which can be interpreted in two dual ways:
first, these languages are extremely simple, which means that one could have
hoped for using directly another well-quasi-ordered set without resorting to
finite words; second, any time one encounters a \kl{well-quasi-ordered}
language, it is going to provide relatively efficient algorithms.
{
\renewcommand{\labelenumi}{R\arabic{enumi}}
\begin{enumerate}
    \setcounter{enumi}{2}
    \item All languages $L \subseteq \Sigma^*$ are 
        \kl{well-quasi-ordered} by the \kl{prefix relation} 
        (resp. \kl{suffix relation}) have an \kl{ordinal height} of at most $\omegaOrd$,
        a finite \kl{ordinal width}, and a \kl{maximal order type} strictly
        below $\omegaOrd^2$.
        \hfill (\cref{prefixes-ordinal-invariants:cor} p.\pageref{prefixes-ordinal-invariants:cor})
\end{enumerate}
}

\AP The straightforward generalization of the results for the \kl{prefix
relation} and the \kl{suffix relation} to the \kl{infix relation} is not
possible. Indeed, it follows from Kuske's result that $\Sigma^*$ equipped with
the \kl{subword relation} can be embedded into $\set{a,b}^*$ with the \kl{infix
relation}~\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}. This implies that there
are \kl{well-quasi-ordered} languages for the \kl{infix relation} that have
very large \kl{ordinal invariants}: for instance, the \kl{maximal order type}
of the \kl{subword relation} is $\omegaOrd^{\omegaOrd^{\card{\Sigma} - 1}}$, which
equals its \kl{ordinal width} \cite[Corollary 3.9, Theorem 4.21]{DZSCSC20}.
However, not all is lost if we restrict ourselves to \kl{bounded languages},
i.e., languages that are included in some $w_1^* \cdots w_k^*$ for some finite
choice of words $w_1, \ldots, w_k$. We are able to show the following:

{
\renewcommand{\labelenumi}{R\arabic{enumi}}
\begin{enumerate}
	\setcounter{enumi}{3}
    \item A \kl{bounded language} $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is (included in) a finite union of languages $S_i \cdot P_i$, where each $S_i$ 
        is a \kl{chain} for the \kl{suffix relation},
        and where
        each $P_i$ is a \kl{chain} for the \kl{prefix relation}.

        \hfill
        (\cref{bounded-language:thm} p.\pageref{bounded-language:thm})
\end{enumerate}
}

This result directly translates into upper bounds on the possible \kl{ordinal invariants}
of such languages similarly as for the \kl{prefix relation}. Notice that these
upper bounds are significantly smaller than \kl{ordinal invariants} of the \kl{subword
relation}.
{
\renewcommand{\labelenumi}{R\arabic{enumi}}
\begin{enumerate}
    \setcounter{enumi}{4}
    \item All \kl{bounded languages}
            $L \subseteq \Sigma^*$ that are \kl{well-quasi-ordered} by the \kl{infix relation}
            have an \kl{ordinal height} of at most $\omega$,
            an \kl{ordinal width} strictly below $\omega^2$, and
            a \kl{maximal order type} strictly below $\omega^3$.
\end{enumerate}
}

\AP Finally, we provide a decision procedure for checking whether a language is
\kl{well-quasi-ordered}, which takes as inputs languages recognized by
\kl{amalgamation systems} \cite{ASZZ24}. Such systems will be formally
introduced in
\cref{amalgamation-systems:subsec}, but
for the moment let us just say that they include many classical computational
models such as finite automata, context-free grammars, and Petri nets
\cite{ASZZ24}. This provides us with a \emph{meta-algorithm} for deciding
whether a given language is \kl{well-quasi-ordered} by the \kl{prefix
relation}, the \kl{suffix relation}, or the \kl{infix relation} under mild
effective restrictions on the computational model that we call an \kl{effective
amalgamative class}, the formal definition of which we defer to
\cref{infixes-amalgamation-effective:subsec}.

{
\renewcommand{\labelenumi}{R\arabic{enumi}}
\begin{enumerate}
    \setcounter{enumi}{6}
    \item 
        Let $L$ be a language recognized by an \kl{amalgamation system}.
        If $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
        then $L$ is a \kl{bounded language}.

        \hfill (\cref{infix-amalgamation:thm} p.\pageref{infix-amalgamation:thm})
    \item 
        Let $\mathcal{C}$ be a \kl{strongly effective amalgamative class} of languages.
        There is a decision procedure that takes as input a language $L \in \mathcal{C}$,
        and decides whether $L$ is \kl{well-quasi-ordered} by the \kl{prefix relation},
        the \kl{suffix relation}, or the \kl{infix relation}. 
        \hfill (\cref{infix-wqo-is-emptiness:thm} p.\pageref{infix-wqo-is-emptiness:thm})
\end{enumerate}
}
Let us point out that the above result implies that the hypothesis of
\kl{bounded languages} on the theoretical side is not a restriction in
practice. As a down-to-earth and more easily understandable 
consequence,
our generic decision procedure applies 
to the class $\mathcal{C}_\text{aut}$ of all languages recognized by finite automata,
and to the class $\mathcal{C}_\text{cfg}$ of all languages recognized by context-free grammars,
which are both \kl{effective amalgamative classes} \cref{aut-cfg-infix:cor}.

@@

\subparagraph{Outline} 
We introduce in \cref{prelims:sec} the
necessary background on \kl{well-quasi-orders} and \kl{ordinal invariants}.
In
\cref{prefixes:sec}, which is relatively
self-contained, we study the \kl{prefix relation} and prove in
\cref{prefixes:thm} the characterization of \kl{well-quasi-ordered}
languages by the \kl{prefix relation}. In
\cref{infixes-regular:sec}, we
obtain the \kl[infix relation]{infix} analogue of \cref{prefixes:thm}
specifically for \kl{bounded languages}
(\cref{bounded-language:thm}). Finally, 
we generalize this result to all
\kl{amalgamation systems} in \cref{infixes-amalgamation:sec}
in
(\cref{infix-amalgamation:thm}),
and provide a decision procedure for checking whether a language is
\kl{well-quasi-ordered} by the \kl{infix relation} in
this context (\cref{infix-wqo-is-emptiness:thm}).

\subparagraph{Acknowledgements} We would like to thank participants of the 2024
edition of \kl{Autobóz} for their helpful comments and discussions.

>>>>

\AP When considering an algorithm based on a \kl{well-quasi-ordering}, the
runtime of the algorithm is deeply related to the ``complexity'' of the
underlying quasi-order~\cite{SCHMITZ17}. One way to measure this complexity is
to consider its so-called \kl{ordinal invariants}: for instance, the
\kl{maximal order type} (or \kl{m.o.t.}), originally defined by De Jongh and Parikh
\cite{dejongh77}, is the order type of the maximal linearization of a
well-quasi-ordered set. In the case of a finite set, the \kl{m.o.t.} is precisely
the size of the set. Better runtime bounds were obtained by considering two
other parameters~\cite{SCHMITZ19}: the \kl{ordinal height} introduced by
Schmidt \cite{schmidt81}, and the \kl{ordinal width} of Kříž and Thomas
\cite{kriz90b}. Therefore, when characterizing \kl{well-quasi-ordered}
languages, we will also be interested in deriving upper bounds on their
\kl{ordinal invariants}. We refer to \cref{ordinal-invariants:subsec} for a
more detailed introduction to these parameters and ordinal computations in
general.

\subparagraph{Contribution} In this paper, we focus on words over a finite
alphabet $\Sigma$, and characterize subsets $L \subseteq \Sigma^*$ that are
\kl{well-quasi-ordered} by the \kl{prefix relation}, the \kl{suffix relation},
and the \kl{infix relation}. Furthermore, we devise decision algorithms
whenever the languages are given by reasonable computational models. Finally,
we provide precise bounds on the possible \kl{ordinal invariants} of such
languages.

In the case of the \kl{prefix} and \kl{suffix} relations, we show that a
language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{prefix
relation} (resp. \kl{suffix}) if and only if it is a finite union of
\kl{chains} of the \kl{prefix relation} (resp. \kl{prefix}), where a
\intro{chain} is a totally ordered set that is well-founded. Note that
\kl{chains} are the simplest possible \kl{well-quasi-ordered} sets (they are
\kl{totally ordered} and \kl{well-founded}) and it is known that finite unions
of \kl{well-quasi-ordered} sets are \kl{well-quasi-ordered}. As a consequence,
the above characterization states that only the simplest possible
\kl{well-quasi-ordered} sets are \kl{well-quasi-ordered} by the \kl{prefix
relation} (\cref{prefixes:thm}). Let us furthermore highlight that this
characterization holds without any restriction on the decidability of the
language $L$ itself, but heavily relies on the assumption that $\Sigma$ is
finite. This allows us to derive tight bounds on the \kl{ordinal invariants} of
such \kl{well-quasi-ordered} languages, which can be interpreted in two dual
ways: first, these languages are extremely simple, which means that one could
have hoped for using directly another well-quasi-ordered set without resorting
to finite words; second, any time one encounters a \kl{well-quasi-ordered}
language, it is going to provide relatively efficient algorithms. Indeed, we
prove that all languages $L \subseteq \Sigma^*$ are \kl{well-quasi-ordered} by
the \kl{prefix relation} (resp. \kl{suffix relation}) have an \kl{ordinal
height} of at most $\omegaOrd$, a finite \kl{ordinal width}, and a \kl{maximal
order type} strictly below $\omegaOrd^2$
(\cref{prefixes-ordinal-invariants:cor}).

\AP The straightforward generalization of the results for the \kl{prefix
relation} and the \kl{suffix relation} to the \kl{infix relation} is not
possible. Indeed, it follows from Kuske's result that $\Sigma^*$ equipped with
the \kl{subword relation} can be embedded into $\set{a,b}^*$ with the \kl{infix
relation}~\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}. This implies that there
are \kl{well-quasi-ordered} languages for the \kl{infix relation} that have
very large \kl{ordinal invariants}: for instance, the \kl{maximal order type}
of the \kl{subword relation} is $\omegaOrd^{\omegaOrd^{\card{\Sigma} - 1}}$,
which equals its \kl{ordinal width} \cite[Corollary 3.9, Theorem
4.21]{DZSCSC20}. We show that in two situations, this can be avoided: when the
language is \kl{downwards closed} (i.e., when it is closed under taking
\kl{infixes}), and when the language is \kl(language){bounded} (i.e., 
when it is included in some $w_1^* \cdots w_k^*$ for some finite choice of
words $w_1, \ldots, w_k$). In those cases, we are able to characterize
\kl{well-quasi-ordered} languages by the \kl{infix relation} and derive tight
bounds on their \kl{ordinal invariants}.

In the case of \kl{bounded languages}, we prove that a \kl{bounded language} $L
\subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if
and only if it is (included in) a finite union of languages $S_i \cdot P_i$,
where each $S_i$ is a \kl{chain} for the \kl{suffix relation}, and where each
$P_i$ is a \kl{chain} for the \kl{prefix relation}
(\cref{bounded-language:thm}) This result directly translates into upper bounds
on the possible \kl{ordinal invariants} of such languages similarly as for the
\kl{prefix relation}. Notice that these upper bounds are significantly smaller
than \kl{ordinal invariants} of the \kl{subword relation}: they have an \kl{ordinal
height} of at most $\omegaOrd$, an \kl{ordinal width} strictly below
$\omegaOrd^2$, and a \kl{maximal order type} strictly below $\omegaOrd^3$
(\cref{ordinal-invariants-bounded:cor}).

In the case of \kl{downwards closed} languages, we prove that they are deeply
related to the notion of \kl{uniformly recurrent words}, borrowed from the
study of word combinatorics. As an intermediate result, we prove that an
infinite word $w$ is \kl{ultimately uniformly recurrent} if and only if the set
of all \kl{infixes} of $w$ is \kl{well-quasi-ordered} by the \kl{infix
relation} (\cref{ultimately-uniformly-recurrent:lem}). We show that every
language $L \subseteq \Sigma^*$ that is \kl{downwards closed} and
\kl{well-quasi-ordered} by the \kl{infix relation} is a finite union of the
sets of finite \kl{infixes} of some \kl{ultimately uniformly recurrent}
bi-infinite words. This also proves that such languages have an \kl{ordinal
height} of at most $\omegaOrd$, an \kl{ordinal width} strictly below
$\omegaOrd^2$, and a \kl{maximal order type} strictly below $\omegaOrd^3$
(\cref{small-ordinal-invariants:thm}).

Then, we turned our attention to decision procedures. To that end, we need to
choose a computational model representing languages. For \kl{downwards closed}
languages, because of their close connection with infinite words, we considered
a model based on \kl{automatic sequences}. Using this model, we can decide
whether a language is \kl{well-quasi-ordered} by the \kl{infix relation}
(\cref{automatic-wqo:thm}). We also studied another representation, where
languages are recognized by \kl{amalgamation systems} \cite{ASZZ24}. Such
systems will be formally introduced in \cref{amalgamation-systems:subsec}, but
for the moment let us just say that they include many classical computational
models such as finite automata, context-free grammars, and Petri nets
\cite{ASZZ24}. This provides us with a \emph{meta-algorithm} for deciding
whether a given language is \kl{well-quasi-ordered} by the \kl{prefix
relation}, the \kl{suffix relation}, or the \kl{infix relation} under mild
effective restrictions on the computational model that we call an \kl{effective
amalgamative class}, the formal definition of which we defer to
\cref{infixes-amalgamation-effective:subsec}. Given a class $\mathcal{C}$ that
is a \kl{strongly effective amalgamative class} of languages, we designed a
decision procedure that takes as input a language $L \in \mathcal{C}$, and
decides whether $L$ is \kl{well-quasi-ordered} by the \kl{prefix relation}, the
\kl{suffix relation}, or the \kl{infix relation}
(\cref{infix-wqo-is-emptiness:thm}). Quite surprisingly, we also showed that,
if a language recognized by an \kl{amalgamation system} is
\kl{well-quasi-ordered} for the \kl{infix relation}, then it is a \kl{bounded
language} (\cref{infix-amalgamation:thm}), which automatically bounds the
\kl{ordinal invariants} of the language. Let us point out that the above result
implies that the hypothesis of \kl{bounded languages} on the theoretical side
is not a restriction in practice. As a down-to-earth and more easily
understandable consequence, our generic decision procedure applies to the class
$\mathcal{C}_\text{aut}$ of all languages recognized by finite automata, and to
the class $\mathcal{C}_\text{cfg}$ of all languages recognized by context-free
grammars, which are both \kl{effective amalgamative classes}
(\cref{aut-cfg-infix:cor}).

Finally, we noticed that for \kl{downwards closed} languages that are
\kl{well-quasi-ordered} by the \kl{infix relation}, being
\kl(language){bounded} is the same as being \kl(language){regular}.
Furthermore, a \kl{bounded language} is \kl{well-quasi-ordered} by the
\kl{infix relation} if and only if its \kl{downwards closure} is
\kl{well-quasi-ordered} by the \kl{infix relation}
(\cref{bounded-wqo-dwclosed:cor}). This shows that, for \kl{bounded languages}
(and therefore, for all languages recognized by \kl{amalgamation systems}) that
are \kl{well-quasi-ordered} by the \kl{infix relation}, their \kl{downwards
closure} is a \kl{regular language}. This is a weak version of the usual result
that the \kl{downwards closure} for the \kl{scattered subword relation} is
always a \kl{regular language}.

@@

\subparagraph{Outline} 
We introduce in \cref{prelims:sec} the
necessary background on \kl{well-quasi-orders} and \kl{ordinal invariants}.
In
\cref{prefixes:sec}, which is relatively
self-contained, we study the \kl{prefix relation} and prove in
\cref{prefixes:thm} the characterization of \kl{well-quasi-ordered}
languages by the \kl{prefix relation}. In
\cref{infixes-bounded:sec}, we
obtain the \kl[infix relation]{infix} analogue of \cref{prefixes:thm}
specifically for \kl{bounded languages}
(\cref{bounded-language:thm}). 
In \cref{infixes-dwclosed:sec}, we study the \kl{downwards closed}
languages, and compute bounds on their \kl{ordinal invariants} in \cref{small-ordinal-invariants:thm}.
Finally, 
we generalize these results to all
\kl{amalgamation systems} in \cref{infixes-amalgamation:sec}
in
(\cref{infix-amalgamation:thm}),
and provide a decision procedure for checking whether a language is
\kl{well-quasi-ordered} by the \kl{infix relation} (resp. \kl{prefix} and \kl{suffix}) in
this context (\cref{infix-wqo-is-emptiness:thm}).

\subparagraph{Acknowledgements} We would like to thank participants of the 2024
edition of \kl{Autobóz} for their helpful comments and discussions.
We would also like to thank Vincent Jugé for his pointers on word combinatorics.

### 2025-01-13 src/prelims.tex
Rework the intro and prelims

\paragraph*{Finite words.} In this paper, we use letters $\Sigma, \Gamma$ to
denote finite alphabets, $\Sigma^*$ to denote the set of finite words over
$\Sigma$, and $\varepsilon$ for the empty word in $\Sigma^*$. In order to give
some intuition on the decision problems, we will sometimes use the notion of
\intro{finite automata}, \intro{regular languages}, and Monadic Second Order
logic ($\intro*\MSO$) over finite words, and assume is familiar with them. We
refer to the textbook of \cite{THOM97} for a detailed introduction to this
topic.

\paragraph*{Orderings and Well-Quasi-Orderings.} A \intro{quasi-order} is a
reflexive and transitive binary relation, it is a \intro{partial order} if it
is furthermore antisymmetric. A \intro{total order} is a \kl{partial order}
where any two elements are comparable. Let now us introduce some notations for
\kl{well-quasi-orders}. A sequence $\seqof{x_i}$ in a set $X$ is
\intro(sequence){good} if there exist $i < j$ such that $x_i \leq x_j$. It is
\intro(sequence){bad} otherwise. Therefore, a \kl{well-quasi-ordered} set is a
set where every infinite sequence is \kl(sequence){good}. A \intro{decreasing
sequence} is a sequence $\seqof{x_i}$ such that $x_{i+1} < x_i$ for all $i$,
and an \intro{antichain} is a set of pairwise incomparable elements. An
equivalent definition of a \kl{well-quasi-ordered} set is that it contains no
infinite \kl{decreasing sequences}, nor infinite \kl{antichains}. We refer to
\cite{SCSC12} for a detailed survey on well-quasi-orders.

@@

\paragraph*{Ordinal Invariants.} \label{ordinal-invariants:subsec}
An \intro{ordinal} is a \kl{well-founded} \kl{totally ordered}
set. We use $\alpha, \beta, \gamma$ to denote ordinals, and use $\intro*\omegaOrd$ to
denote the first infinite \kl{ordinal}, i.e., the set of natural numbers with the
usual ordering. We also use $\intro*\omegaOne$ to denote the first \emph{uncountable}
ordinal.
We only assume superficial familiarity with ordinal arithmetic, and
refer to the books of Kunen \cite{KUNEN80} and Krivine~\cite[Chapter
II]{KRIVINE71} for a detailed introduction to this domain. Given a tree $T$
whose branches are all finite we can define an \kl{ordinal} $\alpha_T$ inductively
as follows: if $T$ is a leaf then $\alpha_T = 0$, if $T$ has children
$\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i} + 1}{i \in \Nat}$. We
say that $\alpha_T$ is the \emph{rank} of $T$.

Let $(X, \leq)$ be a \kl{well-quasi-ordered} set. One can define three
well-founded trees from $X$: the tree of \kl{bad sequences}, the tree of
decreasing sequences, and the tree of \kl{antichains}. The rank of these
respective trees are called respectively the \intro{maximal order type} of $X$
written $\intro*{\oType{X}}$ \cite{dejongh77}, the \intro{ordinal height} of
$X$ written $\intro*{\oHeight{X}}$ \cite{schmidt81}, and the \intro{ordinal
width} of $X$ written $\intro*{\oWidth{X}}$ \cite{kriz90b}. These three
parameters are called the \intro{ordinal invariants} of a
\kl{well-quasi-ordered} set $X$. We refer to the survey of \cite{DZSCSC20} for
a detail discussion on these concepts and their computation on specific classes
of well-quasi-ordered sets.

>>>>

\paragraph*{Finite words.}
\AP 
In this paper, we use letters $\Sigma, \Gamma$ to
denote finite alphabets, $\Sigma^*$ to denote the set of finite words over
$\Sigma$, and $\varepsilon$ for the empty word in $\Sigma^*$. In order to give
some intuition on the decision problems, we will sometimes use the notion of
\intro{finite automata}, \intro{regular languages}, and Monadic Second Order
logic ($\intro*\MSO$) over finite words, and assume is familiar with them. We
refer to the textbook of \cite{THOM97} for a detailed introduction to this
topic.

\paragraph*{Orderings and Well-Quasi-Orderings.}
\AP
A \intro{quasi-order} is a
reflexive and transitive binary relation, it is a \intro{partial order} if it
is furthermore antisymmetric. A \intro{total order} is a \kl{partial order}
where any two elements are comparable. Let now us introduce some notations for
\kl{well-quasi-orders}. A sequence $\seqof{x_i}$ in a set $X$ is
\intro(sequence){good} if there exist $i < j$ such that $x_i \leq x_j$. It is
\intro(sequence){bad} otherwise. Therefore, a \kl{well-quasi-ordered} set is a
set where every infinite sequence is \kl(sequence){good}. A \intro{decreasing
sequence} is a sequence $\seqof{x_i}$ such that $x_{i+1} < x_i$ for all $i$,
and an \intro{antichain} is a set of pairwise incomparable elements. An
equivalent definition of a \kl{well-quasi-ordered} set is that it contains no
infinite \kl{decreasing sequences}, nor infinite \kl{antichains}. We refer to
\cite{SCSC12} for a detailed survey on well-quasi-orders.

@@

\paragraph*{Ordinal Invariants.} \label{ordinal-invariants:subsec}
\AP
An \intro{ordinal} is a \kl{well-founded} \kl{totally ordered}
set. We use $\alpha, \beta, \gamma$ to denote ordinals, and use $\intro*\omegaOrd$ to
denote the first infinite \kl{ordinal}, i.e., the set of natural numbers with the
usual ordering. We also use $\intro*\omegaOne$ to denote the first \emph{uncountable}
ordinal.
We only assume superficial familiarity with ordinal arithmetic, and
refer to the books of Kunen \cite{KUNEN80} and Krivine~\cite[Chapter
II]{KRIVINE71} for a detailed introduction to this domain.
Given a tree $T$
whose branches are all finite we can define an \kl{ordinal} $\alpha_T$ inductively
as follows: if $T$ is a leaf then $\alpha_T = 0$, if $T$ has children
$\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i} + 1}{i \in \Nat}$. We
say that $\alpha_T$ is the \emph{rank} of $T$.

\AP
Let $(X, \leq)$ be a \kl{well-quasi-ordered} set. One can define three
well-founded trees from $X$: the tree of \kl{bad sequences}, the tree of
decreasing sequences, and the tree of \kl{antichains}. The rank of these
respective trees are called respectively the \intro{maximal order type} of $X$
written $\intro*{\oType{X}}$ \cite{dejongh77}, the \intro{ordinal height} of
$X$ written $\intro*{\oHeight{X}}$ \cite{schmidt81}, and the \intro{ordinal
width} of $X$ written $\intro*{\oWidth{X}}$ \cite{kriz90b}. These three
parameters are called the \intro{ordinal invariants} of a
\kl{well-quasi-ordered} set $X$. We refer to the survey of \cite{DZSCSC20} for
a detail discussion on these concepts and their computation on specific classes
of well-quasi-ordered sets.

\AP
We will use the following inequality between \kl{ordinal invariants}, due to
\cite{kriz90b}, and that was recalled in \cite[Theorem 3.8]{DZSCSC20}:
$\oType{X} \leq \oHeight{X} \oComProd \oWidth{X}$, where $\intro*\oComProd$ is
the \intro{commutative ordinal product}, also known as the \reintro{Hessenberg
product}. We will not recall the definition of this product here, and refer to
\cite[Section 3.5]{DZSCSC20} for a detailed introduction to this concept. The
only equalities we will use are $\omegaOrd \oComProd \omegaOrd = \omegaOrd^2$
and $\omegaOrd^2 \oComProd \omegaOrd = \omegaOrd^3$.
